### 2014-04-09 Manuscript/Phase_detectors_notes.tex
First commit - latex notes

\documentclass{article}
\title{Phase paper}
\author{}
\date{}
\begin{document}
\maketitle

\section{Questions}

\begin{itemize}
\item How do you detect phase
\begin{itemize}
\item What are the methods of phase detection
\end{itemize}
\item What can you do with phase once you have it
\item What does phase tell you with respect to physiological systems
\end{itemize}

\section{Problems}

\begin{itemize}
\item noise in data
\item phase slips
\item time-varying frequencies in data
\item choosing how to access the frequencies of interest
\begin{itemize}
\item bandpass and use a range of frequencies
\item methods of choosing the correct frequency at which to acquire the phase
\end{itemize}
\end{itemize}

\section{Methods}
\begin{itemize}
\item Filters
\begin{itemize}
\item filter types and their effects on data
\end{itemize}
\item Phase detectors
\begin{itemize}
\item Hilbert transform
\begin{itemize}
\item what does the HT actually do
\end{itemize}
\item Wavelets
\item Fourier transform
\begin{itemize}
\item is this different from HT and how
\end{itemize}
\end{itemize}
\item Choosing frequency for phase detection
\begin{itemize}
\item averaging frequencies (Hilbert)
\item maximum likelihood method
\item line fitting
\end{itemize}
\end{itemize}

\section{Data}
\begin{itemize}
\item Human
\begin{itemize}
\item BP
\item IBI
\item CBF
\item SCG
\end{itemize}
\item Rat
\begin{itemize}
\item BP
\item RBF
\item Speckle contrast
\end{itemize}
\item Simulation
\begin{itemize}
\item Time series with multiple frequencies corresponding to the systems we are investigating
\begin{itemize}
\item{heart rate}
\item{interbeat interval}
\item{meyer waves?}
\item{myogenic mechanism in kidney}
\item{tubuloglomerular feedback}
\item{respiration?}
\item{others?}
\end{itemize}
\item{set frequencies and adjust them}
\item{add noise to the signal at varying levels}
\item{Make an interaction between two oscillators}
\begin{itemize}
\item{seems like differential equation}
\end{itemize}
\end{itemize}
\end{itemize}
\end{document}